---
title: "Tumer Classification with kNN"
author: "Mohammed Zakaria"
#output:
#  html_document:
#    df_print: paged
#    toc: yes
#    keep_md: true
   
  
#  html_notebook:
#    highlight: espresso
#    theme: readable
#    toc: yes
output: github_document
---


Required packages
```{r message=FALSE, warning= TRUE }
#install.packages("class") # for kNN classification
library(class)
#install.packages("gmodels") # for CrossTable function at the evaluation
library(gmodels)
#install.packages("caret") # for model tuning
library(caret)
#install.packages("e1071") # to help with model tuning
library(e1071)
#install.packages("pROC") # to make ROC plots
library(pROC)	
library(visdat) # better data exploration with vis_dat
library(assertr) # mkae assert statements for better data quality
library(dplyr)
```


Pulling the data: The cancer data is from Brett Lantz's "Machine Learning with R"
a repo for the data is under this link:
https://github.com/mzakariaCERN/Machine-Learning-with-R-datasets/blob/master/wisc_bc_data.csv
and original data can be found under
https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/
```{r}
wbcd <- read.csv(file = "C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/1_kNN/wisc_bc_data.csv", stringsAsFactors = FALSE)


dim(wbcd)
str(wbcd)
summary(wbcd)

# remove id

wbcd <- wbcd[-1]

```


Lets try a better visualization with visdat
```{r}
vis_dat(wbcd)
```

```{r}
vis_miss(wbcd)
```

I'd like to make one asser statement regarding the radius of the tumers

```{r}
invisible(
  wbcd %>% 
    verify(c( "B", "M") %in% .$diagnosis) %>% 
          assert(within_bounds(0,Inf), radius_mean) 
)
```



Convert diagnosis into factors

```{r}
wbcd$diagnosis <- factor(wbcd$diagnosis, levels = c("M", "B"), labels = c("Malignant", "Benign"))
round(prop.table(table(wbcd$diagnosis))*100, digits = 1)

```

Since different features have different scales, we introduce a function "normalize" to set the values withing a range of [0,1]

```{r}
normalize <- function(x){
  return((x - min(x)) / (max(x) - min(x)))
  
  
}
```
use lapply to get a list (the l in lappy) and then convert the list inso data frame

```{r}

wbcd_n <- apply(wbcd[2:31],2,normalize)
#another choice 
#wbcd_n <- as.data.frame(lapply(wbcd[2:31], normalize))



```

We split the data into training and testing (the records wehre already in random order, so no need to randomize further)

```{r}
wbcd_train <- wbcd_n[1:469,]
wbcd_test <- wbcd_n[470:569,]

```

remove the labeling

```{r}
wbcd_train_labels <- wbcd[1:469, 1]
wbcd_test_labels <- wbcd[470:569,1]

```



Building the classifier
```{r}
wbcd_test_pred <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

```



Evaluating model performance 
```{r}
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)

```
Tuning
See what caret has to say about knn
```{r}
modelLookup("knn")


```

Let us do the tuning with some values for k

```{r}
m <- train(wbcd_train,  wbcd_train_labels, method = "knn")
m
```

Something fancier

```{r}
m_cv <- train(wbcd_train, wbcd_train_labels, method = "knn",  trControl = trainControl(method = "cv"), tuneLength = 15) 
# here tuneLength is how many values of k we are going to use
m_boot <- train(wbcd_train, wbcd_train_labels, method = "knn",  trControl = trainControl(method = "boot"), tuneLength = 15)

ctrl <- trainControl(classProbs = TRUE, method = "boot", verboseIter = TRUE)
m_boot_ROC <- train(wbcd_train, wbcd_train_labels, method = "knn",  trControl = ctrl, tuneLength = 15, metric = "ROC")

ctrl <- trainControl(method = "repeatedcv",   # 10fold cross validation
                     number = 5,							# do 5 repititions of cv
                     summaryFunction = twoClassSummary,	# Use AUC to pick the best model
                     classProbs = TRUE,
                     allowParallel = TRUE,
                     verboseIter = TRUE)
m_cv_ROC <- train(wbcd_train, wbcd_train_labels,
      method = "knn",
      metric = "ROC",
      trControl = ctrl)

m_cv

```


To get tons of details about the model and how it was tuned:
```{r}
#str(m_cv)
```

Let us make a prediction
```{r}
m_cv_ROC_prediction <- predict(m_cv_ROC,wbcd_test)
#CrossTable(x = wbcd_test_labels, y = m_cv_ROC_prediction, prop.chisq = FALSE)
confusionMatrix(m_cv_ROC_prediction, wbcd_test_labels)

m_cv_ROC_prediction_probs <- predict(m_cv_ROC,wbcd_test, type = "prob") # you need the prob option to get ROC
#head(m_cv_ROC_prediction_probs)

ROC <- roc(predictor = m_cv_ROC_prediction_probs$Malignant,
               response = wbcd_test_labels,
               levels = rev(levels(wbcd_test_labels)))

ROC$auc

plot(ROC,main = "ROC for kNN")
```