---
title: "Tumer Classification with kNN"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    highlight: espresso
    theme: readable
    toc: yes
---

Required packages
```{r}
#install.packages("class") # for kNN classification
library(class)
#install.packages("gmodels") # for CrossTable function at the evaluation
library(gmodels)
```


```{r}
#plot(cars)
wbcd <- read.csv(file="C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/1_kNN/wisc_bc_data.csv", stringsAsFactors = FALSE)

str(wbcd)
summary(wbcd)

# remove id

wbcd <- wbcd[-1]

```

Convert diagnosis into factors

```{r}
wbcd$diagnosis <- factor(wbcd$diagnosis, levels = c("M", "B"), labels = c("Malignant", "Benign"))
round(prop.table(table(wbcd$diagnosis))*100, digits = 1)

```

Since different features have different scales, we introduce a function "normalize" to set the values withing a range of [0,1]

```{r}
normalize <- function(x){
  return((x - min(x)) / (max(x) - min(x)))
  
  
}
```
use lapply to get a list (the l in lappy) and then convert the list inso data frame

```{r}

wbcd_n <- apply(wbcd[2:31],2,normalize)
#another choice 
#wbcd_n <- as.data.frame(lapply(wbcd[2:31], normalize))



```

We split the data into training and testing (the records wehre already in random order, so no need to randomize further)

```{r}
wbcd_train <- wbcd_n[1:469,]
wbcd_test <- wbcd_n[470:569,]

```

remove the labeling

```{r}
wbcd_train_labels <- wbcd[1:469, 1]
wbcd_test_labels <- wbcd[470:569,1]

```



Building the classifier
```{r}
wbcd_test_pred <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

```



Evaluating model performance 
```{r}
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)

```


