---
title: "R Notebook"
output: github_document
---


This notebook plays a bit with web scraping. 

```{r}
#install.packages("rvest")
library(rvest)
```


```{r}
packt_page <- read_html("https://www.packtpub.com")
packt_page
```

If we want to scrape the papge title, we inspect the output of packt_page and we see there is only one title per page, wrapped withing <title> and <\title> tages.


```{r}
html_node(packt_page, "title")
```

To convert this into plain text, we run the function below:
```{r}
html_node(packt_page, "title") %>% html_text()
```

Another example: Scrap a list of all R-packages on CRAN

```{r}
cran_ml <- read_html("http://cran.r-project.org/web/views/MachineLearning.html")
cran_ml
```

Because we expect more than one result, we use "nodes" instead of "node"

```{r}
ml_packages <- html_nodes(cran_ml, "a")
head(ml_packages, n = 7)
```


## Parsing XML documents
XML is a plain text, human-readable, structured markup language upon which many document formats have been based. It employs a tagging structure in some ways similar to HTML, but is far stricter about formatting. For this reason, it is a popular online format to store structured datasets. 

Recently, the xm12 package has surfaced as an easier and more R-like interface to the libxm12 library. see package github for more details and examples. 


## Parsing JSON from API 
Lets take an example on how to get the altitude and the latitude for the effel tower from google API

```{r}
#install.packages("httr")
library(httr)
map_search <- GET("https://maps.googleapis.com/maps/api/geocode/json",query = list(address = "Effel Tower"))
map_search


```

To access the resulting JSON, which the httr package parsed automatically, we use content() function:

```{r}
content(map_search)
```

To access these contents individually, refer to them using the list syntax:

```{r}
content(map_search)$`results`[[1]]$formatted_address
```

## Converting from JASON to R

```{r}
#install.packages("rjson")
library(rjson)
ml_book <- list(book_title = "Machine Learning with R", author = "Brett Lantz")
toJSON(ml_book)
ml_book_json <- "{\"title\": \"Machine Learning with R\", 
\"Author\": \"Brett Lantz\",
\"publisher\": { 
\"name\": \"Packt Publishing\",
\"url\": \"https://www.packtpub.com\"
},
\"topics\": [\"R\", \"machine learning\", \"data mining\"],
\"MSRP\": 54.99
}"
```

```{r}
ml_book_r <- fromJSON(ml_book_json)
str(ml_book_r)
```

