---
title: "Spam SMS classifier with Naive Bays"
author: "Mohammed Zakaria"
#output:
#  html_document:
#    df_print: paged
#    toc: yes
#    keep_md: true
   
  
#  html_notebook:
#    highlight: espresso
#    theme: readable
#    toc: yes
output: github_document

---



Required packages
```{r}
#install.packages("class") # for kNN classification
library(class)
#install.packages("gmodels") # for CrossTable function at the evaluation
library(gmodels)
#install.packages("caret") # for model tuning
library(caret)
#install.packages("e1071") # to help with model tuning
library(e1071)
#install.packages("pROC") # to make ROC plots
library(pROC)	
```


Pulling the data: The cancer data is from Brett Lantz's "Machine Learning with R"
a repo for the data is under this link:
https://github.com/mzakariaCERN/Machine-Learning-with-R-datasets/blob/master/wisc_bc_data.csv
and original data can be found under
https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/
```{r}
sms_raw <- read.csv(file="C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/2_NB/Data/sms_spam.csv", stringsAsFactors = FALSE)


dim(sms_raw)
str(sms_raw)
summary(sms_raw)

```

We see there are two features. And the feature type has a categorical variables. So we need to convert it to factor


```{r}
sms_raw$type <- as.factor(sms_raw$type)

str(sms_raw$type)
table(sms_raw$type)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
