---
title: "SVM for Optical Character Recoginition"
output: html_notebook
---

```{r load needed libraries}
check.packages("kernlab")
```


```{r loading data}
letters <- read.csv("Data/letterdata.csv")
str(letters)
```

SVM needs all numeric and scaled data. But the package we wil use will do the sacaling for us, and there are no factor data.

```{r split data}
letters_train <- letters[1:16000,]
letters_test  <- letters[16001:20000,]
```

WE will chose kernlab package. It was developed natively in R, and it works well with caret.

```{r build model}
letter_classifier <- ksvm(letter ~ ., data = letters_train, kernel = "vanilladot")
```

```{r}
letter_classifier
```

```{r model evaluation} 
letter_predictions <- predict(letter_classifier, letters_test)
head(letter_predictions)
```


WE compare with the test output using function table

```{r}
table(letter_predictions, letters_test$letter)
```

Convert the matrix into T/F

```{r}
agreement <- letter_predictions == letters_test$letter
table(agreement)
prop.table(table(agreement))
```

WE can try our luck with more complicated kernels. And the first choice is usually Gaussian rbf
```{r model improvement}
letter_classifier_rbf <- ksvm(letter ~ . , data =  letters_train, kernel = "rbfdot")
```

```{r}
letter_predictions_rbf <- predict(letter_classifier_rbf, letters_test)
agreement_rbf <- letter_predictions_rbf == letters_test$letter
table(agreement_rbf)
prop.table(table(agreement_rbf))
```

Let us try some tuning
```{r tuning (takes long time)}
ctrl <- trainControl(method = "boot",   # 10fold cross validation
                     number = 1, # default for resamplingfor each tune twice = 25*2 (takes forever! )
                     #repeats = 1,		    # do 5 repititions of cv
                     #summaryFunction=twoClassSummary,	# Use AUC to pick the best model
                     #classProbs = TRUE, 
                     allowParallel = TRUE,
                     verboseIter = TRUE)
svm.tune <- train(x = letters_train[-1],
                  y = letters_train$letter,
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 4,					# 9 values of the cost function
                  preProc = c("center","scale"),  # Center and scale data
                  #metric="ROC",
                  trControl = ctrl)
```
```{r}
svm.tune
```


```{r}
letter_predictions_tune <- predict(svm.tune, letters_test[-1])
agreement_tune <- letter_predictions_tune == letters_test$letter
table(agreement_tune)
prop.table(table(agreement_tune))
```

