---
title: "Bank Loan classifier with Naive Bays"
author: "Mohammed Zakaria"
#output:
#  html_document:
#    df_print: paged
#    toc: yes
#   keep_md: true
#    always_allow_html: yes

  
#  html_notebook:
#    highlight: espresso
#    theme: readable
#    toc: yes
always_allow_html: yes # this is for wordcloud2 output
output: github_document
---



```{r libraries, message=FALSE, warning= TRUE}

library(ggplot2)
#install.packages("C50") # for DT algorithm
library(C50)
```


Pulling the data: The cancer data is from Brett Lantz's "Machine Learning with R" a repo for the data is under this link: https://github.com/stedy/Machine-Learning-with-R-datasets/blob/master/credit.csv and original data can be found under https://archive.ics.uci.edu/ml

```{r getting data}
credit <- read.csv(file="C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/3_DT/Data/credit.csv", stringsAsFactors = FALSE)
str(credit)

summary(credit)
```

from str() we see that the target feature is actually numerical representing a categorical variable (default vs. no default)
```{r convert target feature into factor}
credit$default <- factor(credit$default)


```


```{r checking few interesting features}
table(credit$checking_balance)

table(credit$default)

```

We divide the data 90:10. WE cannot assume that the data is random. So let us do that

```{r Randomize sample}
set.seed(123)
train_sample <- sample(1000, 900) # get 900 randomly selected numbers, each between 0 and 1000
str(train_sample)
credit_train <- credit[train_sample, ]
credit_test <- credit[-train_sample,]
```


```{r check if the target feature is equally represented}
prop.table(table(credit_train$default))
prop.table(table(credit_test$default))

```

Close! So we can proceed.

```{r Train model with default parameters}
# remove the "default" feature since this is the target one
credit_model <- C5.0(credit_train[-21], credit_train$default)
credit_model

```

