---
title: "Calssifying mushroon with Rule Learner"
author: "Mohammed Zakaria"
#output:
#  html_document:
#    df_print: paged
#    toc: yes
#   keep_md: true
#    always_allow_html: yes

  
#  html_notebook:
#    highlight: espresso
#    theme: readable
#    toc: yes
always_allow_html: yes # this is for wordcloud2 output
output: github_document
---



```{r libraries, message=FALSE, warning= TRUE}
library(gmodels) # for CrossTable
library(ggplot2)
#install.packages("C50") # for DT algorithm
library(C50)
library(caret)
#install.packages('pROC')
library(pROC)
library(dplyr)
library(RWeka) # for 1R learnerrs
## neeed to install java
## then 
#Sys.setenv(JAVA_HOME='C:\\Program Files/Java/jdk-10.0.1/')
#library(rJava) # for 1R learnerrs
```


Pulling the data: The mushroom data is from Brett Lantz's "Machine Learning with R" a repo for the data is under this [link](https://github.com/mzakariaCERN/Machine-Learning-with-R-datasets/blob/master/mushrooms.csv) and original data can be found in the [UCI Repository](https://archive.ics.uci.edu/ml)


```{r getting data}
mushrooms <- read.csv(file = "C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/4_RL/Data/mushrooms.csv", stringsAsFactors = FALSE)
str(mushrooms)

summary(mushrooms)
```

```{r Feature Engineering}
# mushrooms$veil_type has one one level, we might as well drop it
mushrooms$veil_type <- NULL

```




```{r}
table(mushrooms$type)
```

Since we want to develop only rules for this sample (not some unforeseen mushroom type) we will not divide the data into testing and training. 

If we use ZeroR classifier, it will classify all mushrooms as edible (because of the slight majority we saw in mushrooms$type). This is not very useful and we can do better using 1R learnerns. 

