---
title: "Calssifying mushroon with Rule Learner"
author: "Mohammed Zakaria"
#output:
#  html_document:
#    df_print: paged
#    toc: yes
#   keep_md: true
#    always_allow_html: yes

  
#  html_notebook:
#    highlight: espresso
#    theme: readable
#    toc: yes
always_allow_html: yes # this is for wordcloud2 output
output: github_document
---



```{r libraries, message=FALSE, warning= TRUE}
library(gmodels) # for CrossTable
library(ggplot2)
#install.packages("C50") # for DT algorithm
library(C50)
library(caret)
#install.packages('pROC')
library(pROC)
library(dplyr)
library(RWeka) # for 1R and JRip learners
## neeed to install java
## see this video https://www.youtube.com/watch?v=Wp6uS7CmivE
## then 
#Sys.setenv(JAVA_HOME='C:\\Program Files/Java/jdk-10.0.1/')
#.rs.restartR()

library(rJava) # for 1R learnerrs
```


Pulling the data: The mushroom data is from Brett Lantz's "Machine Learning with R" a repo for the data is under this [link](https://github.com/mzakariaCERN/Machine-Learning-with-R-datasets/blob/master/mushrooms.csv) and original data can be found in the [UCI Repository](https://archive.ics.uci.edu/ml)


```{r getting data}
mushrooms <- read.csv(file = "C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/4_RL/Data/mushrooms.csv", stringsAsFactors = FALSE)
str(mushrooms)

summary(mushrooms)
```

We notice two things to be fixed:
1. All data seems to be in string form, so we need to convert that to factors
2. veil_type has only one level, so we better remove it

```{r Feature Engineering}
# just re read the data with setting the strings as factors
mushrooms <- read.csv(file = "C:/Users/mkzak/Documents/GitHub/FunWithR/FunWithR/4_RL/Data/mushrooms.csv")

# mushrooms$veil_type has one one level, we might as well drop it
mushrooms$veil_type <- NULL

```




```{r}
table(mushrooms$type)
```

Since we want to develop only rules for this sample (not some unforeseen mushroom type) we will not divide the data into testing and training. 

If we use ZeroR classifier, it will classify all mushrooms as edible (because of the slight majority we saw in mushrooms$type). This is not very useful and we can do better using 1R learnerns. 


```{r Rule Learning}
mushroom_1R <- OneR(type ~ . , data = mushrooms)
mushroom_1R
```

```{r}
summary(mushroom_1R)
```
 Notice that the classifier accepted 120 poisonous mushrooms as edible! 

```{r Using JRip}
mushroom_JRip <- JRip(type ~ ., data = mushrooms)
mushroom_JRip
```

Notice that the model had no errors using these 9 rules. Also ntice that the rules go as if, else statements

```{r}
summary(mushroom_JRip)
```

